---
title: "Global Disease Burden Analysis"
author: "Bruce Shinga WEMBULUA" # AUTHOR NAMES HERE!
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

This document compares the Disability-Adjusted Life Year (DALY) diseases burden for three countries: COUNTRY 1, COUNTRY 2, and COUNTRY 3. We compare the burden due to communicable, maternal, neonatal, and nutritional diseases (CMNN), non-communicable diseases (NCDs), and overall disease burden for these countries. 

The source data is from the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease (GBD) study. 

```{r child='cmnn_burden.Rmd'}

data_cmnn%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  pivot_wider(names_from = countries, values_from = DALYs)%>%kable()


data_cmnn%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
ggplot()+
  aes(x = year, y= DALYs,color= countries)+
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(x="Years", y="Disability-Adjusted Life Years", color="Countries")

```

```{r child='ncd_burden.Rmd'}

data_ncd%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  pivot_wider(names_from = countries, values_from = DALYs)%>%kable()


DALYs_graph2<-data_ncd%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
ggplot()+
  aes(x = year, y= DALYs,color= countries)+
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(x="Years", y="Disability-Adjusted Life Years", color="Countries")


```

```{r child='overall_burden.Rmd'}

data_overall%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  pivot_wider(names_from = countries, values_from = DALYs)%>%kable()


data_overall%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - All causes - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
ggplot()+
  aes(x = year, y= DALYs,color= countries)+
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(x="Years", y="Disability-Adjusted Life Years", color="Countries")

```


