# Communicable, Maternal, Neonatal, and Nutritional Disease Burden (CMNN)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the communicable diseases data
data_cmnn <- read_csv(here("data", "burden-of-disease-cmnn.csv"))
```

## Table of Estimates for CMNN Burden Over Time

```{r}

# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

DALYs_Table<- data_cmnn%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  pivot_wider(names_from = countries, values_from = DALYs)%>%kable()

DALYs_Table
  

```

## Plot Showing Trends in CMNN Burden Over Time

```{r}

# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

DALYs_graph<-data_cmnn%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Communicable, maternal, neonatal, and nutritional diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
ggplot()+
  aes(x = year, y= DALYs,color= countries)+
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(x="Years", y="Disability-Adjusted Life Years", color="Countries")


```

## Summary of CMNN Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.


- Overall, we observe a decrease in DALYs over time across all three countries.

- However, the trend in DALYs for Côte d'Ivoire is notably distinct from that of Angola and Congo.

- Angola and Congo exhibit a similar trend.




