# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.


View(data_ncd)

data_ncd%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  pivot_wider(names_from = countries, values_from = DALYs)%>%kable()

```

## Plot Showing Trends in NCD Burden Over Time 

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

data_ncd%>%
  select(countries= Entity,year=Year, DALYs=`DALYs (Disability-Adjusted Life Years) - Non-communicable diseases - Sex: Both - Age: Age-standardized (Rate)`)%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
  filter(countries %in% c("Congo", "Angola", "Cote d'Ivoire"))%>%
ggplot()+
  aes(x = year, y= DALYs,color= countries)+
  geom_line()+
  geom_point()+
  theme_minimal()+
  labs(x="Years", y="Disability-Adjusted Life Years", color="Countries")
```

## Summary of NCD Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.


- Overall, we observe a decrease in DALYs over time across all three countries, but notably in Congo and Angola, where DALYs were higher in the 1990s.
